---
title: "Processing a taxatable post MBC and illumina_post_MBC_script2.R"
html_notebook: default
subtitle: ''
classoption: portrait
output: html_document

---

# 12SV51

## load functions

```{r, eval=TRUE, results='hide',echo=F}
bastoolsDir<-"/home/bastian.egeter/git_bastools/bastools/"
setwd(bastoolsDir)
source("master_functions.R")
library(ggplot2)
library(googlesheets4)
sheets_auth(email = "basegeter@gmail.com")
options(httr_oob_default = TRUE)
```

## settings 
```{r, eval=TRUE}
#can be multiple
taxatabs<-c("/mnt/Disk1/BASTIAN_POST_MBC_MISEQS/2020_01/rebins/IRAN-2020_01_none.flash2.vsearch_qfilt.cutadapt.vsearch_uniq.vsearch_afilt.allsamples_step5.ALL_vsearch_uniq.nodenoise.noclust.rebins.taxatable.tf.spliced.txt")

#datasheet
ss_url<-"https://docs.google.com/spreadsheets/d/1Kw8ONKtlX1hGr0LM-FvkowXYElEJ_taFNu2ZSXqp_n8/edit#gid=1531090624"

#options for subsetting master sheet 
subsetlist<-list(experiment_id=c("2020_01"))
```

```{r, eval=TRUE,echo=F}
#settings counts
#files for doing counts (currently processes one run*primer combo only)
 MBCtsvDir<-"/mnt/Disk1/BASTIAN_POST_MBC_MISEQS/2020_01/MBC_OUT_12SV1_X2/"
 otutab<-"/mnt/Disk1/BASTIAN_POST_MBC_MISEQS/2020_01/none.flash2.vsearch_qfilt.cutadapt.vsearch_uniq.vsearch_afilt.allsamples_step5.ALL_vsearch_uniq.nodenoise.noclust.otutab.tsv"
 illumina_script_taxatab<-"/mnt/Disk1/BASTIAN_POST_MBC_MISEQS/2020_01/rebins/2020_01_none.flash2.vsearch_qfilt.cutadapt.vsearch_uniq.vsearch_afilt.allsamples_step5.ALL_vsearch_uniq.nodenoise.noclust.rebins.taxatable.txt"
 illumina_script_taxatab_tf<-"/mnt/Disk1/BASTIAN_POST_MBC_MISEQS/2020_01/rebins/2020_01_none.flash2.vsearch_qfilt.cutadapt.vsearch_uniq.vsearch_afilt.allsamples_step5.ALL_vsearch_uniq.nodenoise.noclust.rebins.taxatable.tf.txt"
```


```{r, eval=TRUE,echo=F}
##organise mastersheet
master_sheet<-google.read.master.url(ss_url)
#remove duplicate columns (should functionise this)
if(length(grep("\\.\\.\\.",colnames(master_sheet),value = T))>0) {
  message("Removing duplicated columns")
  a<-grep("\\.\\.\\.",colnames(master_sheet),value = T)
  b<-do.call(rbind,stringr::str_split(a,"\\.\\.\\."))
  d<-b[duplicated(b[,1]),]
  e<-paste0(d[,1],"...",d[,2])
  master_sheet<-master_sheet[,!colnames(master_sheet) %in% e]
  colnames(master_sheet)<-gsub("\\.\\.\\..*","",colnames(master_sheet))
}

#changing sample_type too to match functions
colnames(master_sheet)<-gsub("Sample_Type","sample_type",colnames(master_sheet))

#check on number of samples in each category
message("Showing xtabs for entire datasheet, including Sample_Type by default")
master_xtabs(master_sheet,columns=c("sample_type",names(subsetlist)))
colnames(master_sheet)
##subset master sheet for this study 
message("Subsetting datasheet")
ms_ss<-subset_mastersheet(master_sheet, subsetlist)
#check again to see the subset made sense
master_xtabs(ms_ss,columns=c("sample_type",names(subsetlist)))
```

# Make counts at various steps of MBC and illumina script
```{r, eval=TRUE,echo=F}
countdf<-count.MBC(MBCtsvDir,ms_ss,otutab,illumina_script_taxatab,illumina_script_taxatab_tf)
```

## import taxatabs
```{r, eval=TRUE,echo=F, results='hide'}
#import taxtabs
all.taxatabs<-bas.merge.taxatabs(taxatabs)

#make new taxa table based on subsetted master sheet
all.taxatabs.ss<-cbind(taxon=all.taxatabs$taxon,all.taxatabs[colnames(all.taxatabs) %in% ms_ss$ss_sample_id])
```

## remove samples/taxa with 0 reads
```{r, eval=TRUE,echo=F}
all.taxatabs.ss<-rm.0readtaxSam(all.taxatabs.ss)
```

## summary ---this is pre-grouping - please inspect, look at groupings and look at post-grouping table. Can more grouping be done, is anything wrong?
```{r, eval=TRUE,echo=F}
DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T,scrollX=T))
```

## group taxa, where possible - Do you both agree with these: e.g. if we have reads at genus level Discoglossus, we can join them to D. galganoi - same for all the joining done below
```{r, eval=TRUE}
# all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon="Eukaryota;Chordata;Amphibia;Anura;Alytidae;Discoglossus;NA",
#                                        jointo="Eukaryota;Chordata;Amphibia;Anura;Alytidae;Discoglossus;Discoglossus galganoi")
# 
# 
# all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon="Eukaryota;Chordata;Amphibia;Anura;Bufonidae;Epidalea;NA",
#                                        jointo="Eukaryota;Chordata;Amphibia;Anura;Bufonidae;Epidalea;Epidalea calamita")
# 
# all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon="Eukaryota;Chordata;Amphibia;Anura;Alytidae;Alytes;NA",
#                                        jointo="Eukaryota;Chordata;Amphibia;Anura;Alytidae;Alytes;Alytes obstetricans")
# 
# all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon="Eukaryota;Chordata;Amphibia;Anura;Pelobatidae;NA;NA",
#                                        jointo="Eukaryota;Chordata;Amphibia;Anura;Pelobatidae;Pelobates;Pelobates cultripes")
```

## summary--post-grouping
```{r, eval=TRUE,echo=F}
DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T,scrollX=T))
```

## apply taxon and sample filters
```{r, eval=TRUE,echo=F}
#Filter samples before removing the NAs and no_hits (0.1%)
all.taxatabs.ss<-taxon.filter.solo.df(taxatab = all.taxatabs.ss,taxonpc = 0.1)
all.taxatabs.ss<-sample.filter.solo(taxatab = all.taxatabs.ss,samplepc=0.1)
```

## remove the NAs and no_hits
```{r, eval=TRUE,echo=F}
all.taxatabs.ss<-all.taxatabs.ss[all.taxatabs.ss$taxon!="NA;NA;NA;NA;NA;NA;NA",]
all.taxatabs.ss<-all.taxatabs.ss[all.taxatabs.ss$taxon!="no_hits;no_hits;no_hits;no_hits;no_hits;no_hits;no_hits",]
countdf$'After project taxon and sample filter'<-sum(all.taxatabs.ss[,-1])
```

## apply detection filter
## check negatives (this is a check only, not removing contaminants at this point). 
## Check removing contaminants (this is a check only, not removing contaminants at this point)
```{r, eval=TRUE,echo=F}
#check with removing detections
message("Applying detection filter")
all.taxatabs.ss.joined.dxn<-filter.dxns(taxatab = all.taxatabs.ss,filter_dxn = 50)

#check negatives
message("Checking negs")
negatives<-negs.stats(taxatab=all.taxatabs.ss.joined.dxn,ms_ss = ms_ss,real=c("faecal_sample"),ex_hominidae=F)

#just to look:
message("Checking effects of contaminant filter")
notprinting<-remove.contaminant.taxa(taxatab = all.taxatabs.ss.joined.dxn,master_sheet = ms_ss,negatives = negatives,group.code = "Sample_Plate")
#note - if I have PCR or extraction negs, need to think about the group carefully

countdf$'After detection filter'<-sum(all.taxatabs.ss.joined.dxn[,-1])

```

## remove contaminants from relevant groups - agree with these decisions? Basically remove Alytes obs then run contaminant script.
```{r, eval=TRUE}
all.taxatabs.ss.joined.dxn.rmcons<-remove.contaminant.taxa(taxatab = all.taxatabs.ss.joined.dxn,master_sheet = ms_ss,negatives = negatives,group.code = "Sample_Plate")

#check negatives again, just to be sure
negatives<-negs.stats(taxatab=all.taxatabs.ss.joined.dxn.rmcons,ms_ss = ms_ss,real=c("faecal_sample"),ex_hominidae=F)

countdf$'After remove contaminants'<-sum(all.taxatabs.ss.joined.dxn.rmcons[,-1])
```

## keep only species-level taxa 
```{r, eval=TRUE,echo=F}
all.taxatabs.ss.joined.dxn.rmcons.spL<-keep.only.spL.assigns(all.taxatabs.ss.joined.dxn.rmcons)
```

## summary and some barplots
```{r, eval=TRUE,echo=F}
#summary
DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss.joined.dxn.rmcons.spL), rownames = FALSE, 
              #filter="top",
          options = list(pageLength = 50, scrollX=T,scrollX=T))

taxatab.stackplot(all.taxatabs.ss.joined.dxn.rmcons.spL,master_sheet = ms_ss,column = "predator",as.percent = F,as.dxns = T)

taxatab.stackplot(all.taxatabs.ss.joined.dxn.rmcons.spL,master_sheet = ms_ss,column = "predator",as.percent = T,as.dxns = F)

taxatab.stackplot(all.taxatabs.ss.joined.dxn.rmcons.spL,master_sheet = ms_ss,column = "biomaterial",as.percent = F,as.dxns = F)

taxatab.stackplot(all.taxatabs.ss.joined.dxn.rmcons.spL,master_sheet = ms_ss,column = "individual",as.percent = F,as.dxns = F,facetcol = "predator")

countdf$'After keeping only species-level taxa'<-sum(all.taxatabs.ss.joined.dxn.rmcons.spL[,-1])
```


## pca plot (just an example, not interesting in this case)
```{r, eval=TRUE,echo=F}
#just an example, not interesting in this case
p<-taxatab.pca.plot.col(taxatab = all.taxatabs.ss.joined.dxn.rmcons.spL,master_sheet = ms_ss,MS_colID = "ss_sample_id",lines = T,longnames = F,shortnames = T,ellipse = T,factor1 = "predator")
p
```

## plot counts
```{r, eval=TRUE,echo=F}
long<-as.data.frame(t(countdf))
long$step<-rownames(long)
long$step<-factor(long$step,levels=long$step)
colnames(long)[1]<-"reads"
p<-ggplot(long,aes(x=step,y=reads,label = long$reads))+geom_bar(stat = "identity") +
   theme(axis.text.x = element_text(angle = 45,hjust=1))+
   theme(axis.line = element_line(colour = "black"))+
   theme(panel.background = element_blank())+
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_continuous(labels = scales::comma)
p

#save plot, outside of rmd
# ggplot2::ggsave(filename = "read.counts.plot.tiff",p,device = "tiff",dpi = "print",)

```

## write final table
```{r, eval=TRUE,echo=F}
#write.table(all.taxatabs.ss.joined.dxn.rmcons.spL,"all.taxatabs.ss.joined.dxn.rmcons.spL.12S1.txt",quote = F, row.names = F,col.names = T,sep = "\t")
```