---
title: "Processing a taxatable post MBC and illumina_post_MBC_script2.R"
html_notebook: default
subtitle: ''
classoption: portrait
output: html_document
---

## read and print settings
```{r read and print settings, eval=TRUE,echo=F}

message("settings:
        ")
message("bastoolsDir: ",bastoolsDir)
message("taxatabs: ",taxatabs)
message("master_sheet_file: ",master_sheet_file)
message("subsetlist: ",subsetlist)
message("config: ",config)
message("out_html: ",out_html)
message("taxatab.out: ",taxatab.out)
message("problemTaxa: ",problemTaxa)
message("fullpc: ",fullpc)
message("taxonpc: ",taxonpc)
message("samplepc: ",samplepc)
message("filter_dxn: ",filter_dxn)
message("filter_dxn2: ",filter_dxn2)
message("real: ",real)
message("neg.types: ",neg.types)
message("neg.groups: ",neg.groups)
message("do.first.negative.inspection: ",do.first.negative.inspection)
message("use.contamination.filter: ",use.contamination.filter)
message("remove.entire.dataset: ",remove.entire.dataset)
message("rm.only.less.than: ",rm.only.less.than)
message("rep.rm: ",rep.rm)
message("rep.rm.first: ",rep.rm.first)
message("rep.rm.second: ",rep.rm.second)
message("xLevel: ",xLevel)
message("zLevel: ",zLevel)
message("aggregate.pre.grouping: ",aggregate.pre.grouping)
message("group.taxa.before.filtering: ",group.taxa.before.filtering)
message("sumrepsby: ",sumrepsby)
message("unwantedTaxa: ",unwantedTaxa)
message("taxa.to.group: ",taxa.to.group)
message("plotting.vars: ",plotting.vars)
message("plotting_levels: ",plotting_levels)
message("hidelegend: ",hidelegend)
message("hidelegend.pca: ",hidelegend.pca)
message("facetcol: ",facetcol)
message("colour.bar: ",colour.bar)
message("krona.out: ",krona.out)
message("plot.bars: ",plot.bars)
message("plot.pca: ",plot.pca)
message("show.pca.lines: ",show.pca.lines)
message("plot.heatmaps: ",plot.heatmaps)

message("Note that 'htf'= higher-than-family throughout")

#print(ls.str())
```


```{r load functions, eval=TRUE, results='hide',echo=F}
source(paste0(bastoolsDir,"master_functions.R"))
library(ggplot2)
library(dplyr)
options(httr_oob_default = TRUE)
all.stats<-list()
stepcounter<-0
```

## organise mastersheet
```{r organise mastersheet, eval=TRUE,echo=F}
if(is.null(master_sheet_file)){
  master_sheet<-google.overlord(url = url,tokenDir = bastoolsDir)
} else {
  master_sheet<-data.table::fread(master_sheet_file,data.table = F,sep = "\t")
}

##subset master sheet for this study 
if(!is.null(subsetlist)) {
  message("Subsetting datasheet")
  ms_ss<-subset_mastersheet(master_sheet, subsetlist)
  #check again to see the subset made sense
  master_xtabs(master_sheet = ms_ss,columns=c("sample_type",names(subsetlist)))
} else ms_ss<-master_sheet
```

## import taxatabs
```{r import taxatabs, eval=TRUE,echo=F,}
#import taxtabs

all.taxatabs<-bas.merge.taxatabs(taxatabs)

#change mbk codes
all.taxatabs$taxon<-gsub("mbk:nb-thr","NA",all.taxatabs$taxon)

#make new taxa table based on subsetted master sheet
all.taxatabs.ss<-cbind(taxon=all.taxatabs$taxon,all.taxatabs[colnames(all.taxatabs) %in% ms_ss$ss_sample_id])
if(length(colnames(all.taxatabs.ss))==1) stop("No matching ss_sample_ids from master sheet found in taxatable")

#remove 0 reads
all.taxatabs.ss<-rm.0readtaxSam(all.taxatabs.ss)

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(taxatab = all.taxatabs.ss,stepname = "start")

#rank stats
print(stats.by.rank(taxatab = all.taxatabs.ss))

#show included/excluded
#message("fix this")
 message("Samples in final dataset")
 print(colnames(all.taxatabs.ss[,-1]))
 message("Samples excluded from final dataset")
 print(master_sheet$ss_sample_id[! master_sheet$ss_sample_id %in% colnames(all.taxatabs.ss[,-1])])
```

# qplots
```{r first qplot, eval=TRUE,echo=F}
  plot1<-qplot.taxatab(all.taxatabs.ss,rm.nohits = T,rm.NA = T,specReadCount = filter_dxn)
  print(plot1)
```

## remove problem taxa - usually human, predator, NAs and no_hits
```{r remove problem taxa, eval=TRUE,echo=F,fig.height=8}
message("Removing no_hits")
all.taxatabs.ss<-all.taxatabs.ss[all.taxatabs.ss$taxon!="no_hits;no_hits;no_hits;no_hits;no_hits;no_hits;no_hits",]

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "rm.nohits")

message("Removing NAs - sequences that had blast hits but were not assigned to any taxon")
#replace "unknown" with NA (these come from database entries being "unknown" at certain level) 
all.taxatabs.ss$taxon<-gsub(";unknown;",";NA;",all.taxatabs.ss$taxon)
all.taxatabs.ss$taxon<-gsub(";unknown;",";NA;",all.taxatabs.ss$taxon) #need to repeat to find all patterns

#sum taxa that are now duplicates
all.taxatabs.ss<-aggregate(all.taxatabs.ss[,-1,drop=F],by = list(taxon=all.taxatabs.ss$taxon),FUN=sum)

all.taxatabs.ss<-all.taxatabs.ss[all.taxatabs.ss$taxon!="NA;NA;NA;NA;NA;NA;NA",]
all.taxatabs.ss<-all.taxatabs.ss[all.taxatabs.ss$taxon!="unknown;unknown;unknown;unknown;NA;NA;NA",]

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "rm.non-assigned")

message("Removing problem taxa")
all.taxatabs.ss<-rm.taxa(taxatab = all.taxatabs.ss,taxaGrep = problemTaxa)

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "rm.problem")

#rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))

if(do.first.negative.inspection==T) full.negative.inspection(taxatab=all.taxatabs.ss,ms_ss=ms_ss,real=real)

```
## group taxa only done here if group.taxa.before.filtering=T
```{r group taxa before filtering, eval=TRUE,echo=F}
if(group.taxa.before.filtering==T){
  if(is.null(taxa.to.group)){
    message("Skipping grouping of taxa")
  } else {
  
    message("Pre-grouping")
    DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T))
    
    #rank stats
    print(stats.by.rank(taxatab = all.taxatabs.ss))
    
    message("Taxa to group")
    print(taxa.to.group)
    
    for(i in 1:nrow(taxa.to.group)){
      all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon=as.character(taxa.to.group[i,1]), jointo=as.character(taxa.to.group[i,2]))
    }
    
    message("Post-grouping")
    DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T))
    
    #rank stats
    print(stats.by.rank(taxatab = all.taxatabs.ss))
    
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "group_taxa")
  }
}
```

## apply taxon and sample filters
```{r apply taxon and sample filters, eval=TRUE,echo=F}
if(taxonpc>0 | samplepc>0 | fullpc>0) {
  
  #apply overall filter
  all.taxatabs.ss<-taxatab.filter.full(all.taxatabs.ss,pc=fullpc)
  stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "fullpc")
  
    all.taxatabs.ss<-taxon.filter.solo.df(taxatab = all.taxatabs.ss,taxonpc = taxonpc)
    
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "taxonpc")
    
    all.taxatabs.ss<-sample.filter.solo(taxatab = all.taxatabs.ss,samplepc=samplepc)
    
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "samplepc")
    
    full.negative.inspection(taxatab=all.taxatabs.ss,ms_ss=ms_ss,real=real)
    
    #rank stats
    print(stats.by.rank(taxatab = all.taxatabs.ss))
  
}else message("taxonpc and samplepc set to 0...skipping taxon and sample filters")



```


## remove detection in less than 2 reps (only done here if rep.rm.first is TRUE)
### PCR negatives are exempt, if the negative option is set (only makes sense pre-rm.contaminants function, or without rm.contaminants function)
```{r remove detection in less than 2 reps rep.rm.first TRUE, eval=TRUE,echo=F}
if (rep.rm.first==T){
  negatives<-negs.stats(taxatab=all.taxatabs.ss,ms_ss = ms_ss,real=real,ex_hominidae=F,printnegs = F)
  all.taxatabs.ss<-rm.single.rep.dxn(taxatab = all.taxatabs.ss,ms_ss = ms_ss,grouping = rep.rm,negatives = negatives)
  
  stepcounter<-stepcounter+1
  all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "repfilter1")

  #full.negative.inspection(taxatab=all.taxatabs.ss,ms_ss=ms_ss,real=real)
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
  
} else message("rep.rm.first=F...skipping removal of detections in less than 2 reps")
```

  
## apply detection filter
```{r apply detection filter, eval=TRUE,echo=F}
#check with removing detections
if(filter_dxn>0){

  message("Applying detection filter")
  all.taxatabs.ss<-filter.dxns(taxatab = all.taxatabs.ss,filter_dxn = filter_dxn)
  
  stepcounter<-stepcounter+1
  all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "dxnfilter")
  
  full.negative.inspection(taxatab=all.taxatabs.ss,ms_ss=ms_ss,real=real)
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
  
  # qplots
  message("qplot after detection filter")
  plot1<-qplot.taxatab(all.taxatabs.ss,rm.nohits = T,rm.NA = T,specReadCount = filter_dxn)
  print(plot1)
  
} else message("filter_dxn=0...skipping detection filter")

```

## remove contaminants 
```{r remove contaminants , eval=TRUE,echo=F}

#use.contamination.filter function?
if(use.contamination.filter==T){
  
  negatives<-suppressMessages(negs.stats(taxatab = all.taxatabs.ss,ms_ss = ms_ss,real = real,ex_hominidae = F,printnegs = F))

  if(is.null(negatives)) {
    message("No negatives with reads, skipping contaminant filter")
    } else {
  
      message("Applying contaminant filter")
      group.codes2<-list()
      for(i in 1:length(negatives)){
      group.codes2[[i]]<-neg.groups[match(names(negatives)[i],neg.types)]
      }
      group.codes2<-do.call(c,group.codes2)
      
      all.taxatabs.ss<-remove.contaminant.taxa(taxatab = all.taxatabs.ss,master_sheet = ms_ss,negatives = negatives,
                                               group.codes = group.codes2,remove.entire.dataset = remove.entire.dataset,
                                               rm.only.less.than=rm.only.less.than,printcontaminations = T)
      
      #check negatives again, just to be sure
      message("Checking that all negatives are now clean")
      
      full.negative.inspection(taxatab=all.taxatabs.ss,ms_ss=ms_ss,real=real)
      
      stepcounter<-stepcounter+1
      all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "rm.contaminants")
  
      #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
  
      } 
} else {
  message("use.contamination.filter set to FALSE, skipping contaminant filter")
}
```

## remove detection in less than 2 reps 
### PCR negatives are exempt, if the negative option is set (only makes sense pre-rm.contaminants function, or without rm.contaminants function)
```{r remove detection in less than 2 reps, eval=TRUE,echo=F}
if (rep.rm.second==T){
  all.taxatabs.ss<-rm.single.rep.dxn(taxatab = all.taxatabs.ss,ms_ss = ms_ss,grouping = rep.rm)
  
  stepcounter<-stepcounter+1
  all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "repfilter2")
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
  
} else message("rep.rm.second=F...skipping removal of detections in less than 2 reps")

```

## sum reps
```{r sum reps, eval=TRUE,echo=F}
all.taxatabs.ss<-sumreps(taxatab = all.taxatabs.ss,ms_ss = ms_ss,grouping = sumrepsby,discard = F)

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(taxatab = all.taxatabs.ss,stepname = "sumreps")

#rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
```

## apply 2nd detection filter (possibly makes more sense here?)
```{r apply 2nd detection filter, eval=TRUE,echo=F}
if(filter_dxn2>0){
  message("Applying 2nd detection filter")
  all.taxatabs.ss<-filter.dxns(taxatab = all.taxatabs.ss,filter_dxn = filter_dxn2)
  
  stepcounter<-stepcounter+1
  all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "dxnfilter2")
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
  
} else message("filter_dxn2=0...skipping 2nd detection filter")
```

## aggregate at chosen level and keep only that-level taxa 
```{r aggregate at chosen level and keep only that-level taxa Pre-grouping, eval=TRUE,echo=F}
if(aggregate.pre.grouping==T){
  if(is.null(xLevel)) {
    message("Skipping aggregate of taxa at xLevel")
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "aggregate_by_xLevel") 
    
  } else {
    all.taxatabs.ss<-aggregate.at.xLevel(taxatab=all.taxatabs.ss,xLevel = xLevel)
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "aggregate_by_xLevel") 
  }
    
  if(is.null(zLevel)) {
    message("Skipping removal of taxa at zlevel")
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "keep_only_zLevel") 
  } else {
    all.taxatabs.ss<-keep.below.xLevel.assigns(taxatab = all.taxatabs.ss,xLevel = zLevel)
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "keep_only_zLevel") 
  }
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
}
```
                                                        

## remove unwanted taxa for analysis
```{r remove unwanted taxa for analysis, eval=TRUE,echo=F}
message("Removing unwanted taxa")
all.taxatabs.ss<-rm.taxa(taxatab = all.taxatabs.ss,taxa = unwantedTaxa)

stepcounter<-stepcounter+1
all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "unwanted")

#rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
```

## group taxa, where possible (only done here if group.taxa.before.filtering=F and taxa.to.group!=NULL)
```{r group taxa, eval=TRUE,echo=F}
if(group.taxa.before.filtering==F){
  if(is.null(taxa.to.group)){
    message("Skipping grouping of taxa")
  } else {
    
    message("Pre-grouping")
    DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T))
    
    #rank stats
    print(stats.by.rank(taxatab = all.taxatabs.ss))
    
    message("Taxa to group")
    print(taxa.to.group)
    
    for(i in 1:nrow(taxa.to.group)){
      all.taxatabs.ss<-bas.group.taxa(taxatab = all.taxatabs.ss,taxon=as.character(taxa.to.group[i,1]), jointo=as.character(taxa.to.group[i,2]))
    }
    
    message("Post-grouping")
    DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T))
    
    #rank stats
    print(stats.by.rank(taxatab = all.taxatabs.ss))
    
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "group_taxa")
  }
}
```

## aggregate at chosen xlevel and keep only zlevel taxa 
```{r aggregate at chosen level and keep only that-level taxa Post-grouping, eval=TRUE,echo=F}
if(aggregate.pre.grouping==F){
  if(is.null(xLevel)) {
    message("Skipping aggregate of taxa at xLevel")
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "aggregate_by_xLevel") 
    
  } else {
    all.taxatabs.ss<-aggregate.at.xLevel(taxatab=all.taxatabs.ss,xLevel = xLevel)
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "aggregate_by_xLevel") 
  }
    
  if(is.null(zLevel)) {
    message("Skipping removal of taxa at zlevel")
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "keep_only_zLevel") 
  } else {
    all.taxatabs.ss<-keep.below.xLevel.assigns(taxatab = all.taxatabs.ss,xLevel = zLevel)
    stepcounter<-stepcounter+1
    all.stats[[stepcounter]]<-taxatab.sumStats(all.taxatabs.ss,stepname = "keep_only_zLevel") 
  }
  
  #rank stats
  print(stats.by.rank(taxatab = all.taxatabs.ss))
}
```

## some barplots
```{r barplots, eval=TRUE,echo=F,fig.height=7,fig.width=20}
if(!is.null(plotting.vars)){
#this multilevel list used for subsequent plots also
taxatabs.multilevels<-list()
  
  if(!is.null(plotting_levels)){
    for(i in 1:length(plotting_levels)){
      #message("Plotting at ",plotting_levels[i]," level")
      if(plotting_levels[i]=="species") {
        taxatabs.multilevels[[i]]<-keep.below.xLevel.assigns(all.taxatabs.ss,xLevel = "species",rm.trailing.NA = F)
        names(taxatabs.multilevels)[i]<-"species"
      } else  taxatabs.multilevels[[i]]<-aggregate.at.xLevel(taxatab = all.taxatabs.ss,
                                         xLevel = plotting_levels[i],rm.above = T,rm.trailing.NA=F)
      names(taxatabs.multilevels)[i]<-plotting_levels[i]
    }
    #taxatabs.multilevels[[length(plotting_levels)+1]]<-all.taxatabs.ss
    #names(taxatabs.multilevels)[length(plotting_levels)+1]<-"combined"
    
    message("Note that taxa paths in plots will be collapsed based on the plotting_level selected. E.g. 'Euk.Cho.Ave.Pro.col.col.col_NA'")
    
  } else taxatabs.multilevels[[1]]<-all.taxatabs.ss
}


if(plot.bars) if(!is.null(plotting.vars)){

  message("Making barplots")
    
  if(!exists("plot.as.percent")) plot.as.percent<-c(F,T)
  if(!exists("plot.as.dxns")) plot.as.dxns<-c(F,T)
  
  if(is.null(facetcol)){
    
    for(h in 1:length(plot.as.dxns)){
      for(k in 1:length(plot.as.percent)){
        for(j in 1:length(taxatabs.multilevels)){
          for(i in 1:length(plotting.vars)){
            print(taxatab.stackplot(taxatab = taxatabs.multilevels[[j]],master_sheet = ms_ss,column = plotting.vars[i],
                                   as.percent = plot.as.percent[k],as.dxns = plot.as.dxns[h], 
                                   hidelegend = hidelegend,grouping = sumrepsby))
                                  
         }
       }
     }
   }
  } else {
      for(h in 1:length(plot.as.dxns)){
        for(k in 1:length(plot.as.percent)){
          for(j in 1:length(taxatabs.multilevels)){
            for(i in 1:length(plotting.vars)){
              if(plotting.vars[i]!=facetcol){
                print(taxatab.stackplot(taxatab = taxatabs.multilevels[[j]],master_sheet = ms_ss,
                                      column = plotting.vars[i],as.percent = plot.as.percent[k],
                                  as.dxns = plot.as.dxns[h], hidelegend = hidelegend,
                                  grouping = sumrepsby,facetcol = facetcol))
              }
            
          }
        }
      }
    }
  }
      
} else message("plotting.vars=F...skipping barplots")
```

## ordered barplots
```{r ordered_barplots, eval=TRUE,echo=F,fig.height=7,fig.width=20}
  #ordered by reads
  all.read.count<-data.frame(taxon=all.taxatabs.ss$taxon,reads=rowSums(all.taxatabs.ss[,-1,drop=F]))
  all.read.count<-all.read.count[order(all.read.count$reads,decreasing = T),]
  all.read.count<-all.read.count[1:50,]
  # all.read.count$taxon<-factor(all.read.count$taxon,ordered = T)
  ggplot(data=all.read.count,aes(x=reorder(taxon,-reads),y=reads))+geom_bar(stat = "identity")+
  theme(legend.title = element_text(size=8), legend.text=element_text(size=8),
          axis.text.x=element_text(size=8,angle=45, hjust=1),legend.position="right",legend.direction="vertical",
          title = element_text(size=6),axis.title = element_text(size=8),panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #ggtitle(paste("x axis=",column,"; as.percent=",as.percent,"; as.dxns=",as.dxns, ";facetcol=",facetcol),subtitle = )+
    xlab("Read counts for top 50 taxa") + scale_y_continuous(labels = scales::comma) +
    ylab("\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\nreads")
  
  #ordered by samples
  all.sample.count<-binarise.taxatab(all.taxatabs.ss,t = T)
  all.sample.count<-data.frame(taxon=all.sample.count$taxon,samples=rowSums(all.sample.count[,-1,drop=F]))
  all.sample.count<-all.sample.count[order(all.sample.count$samples,decreasing = T),]
  all.sample.count<-all.sample.count[1:50,]
  ggplot(data=all.sample.count,aes(x=reorder(taxon,-samples),y=samples))+geom_bar(stat = "identity")+
  theme(legend.title = element_text(size=8), legend.text=element_text(size=8),
          axis.text.x=element_text(size=8,angle=45, hjust=1),legend.position="right",legend.direction="vertical",
          title = element_text(size=6),axis.title = element_text(size=8),panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #ggtitle(paste("x axis=",column,"; as.percent=",as.percent,"; as.dxns=",as.dxns, ";facetcol=",facetcol),subtitle = )+
    xlab(paste("No. of samples that top 50 taxa occurred in. Total no. samples (with reads remaining)=",
         ncol(all.taxatabs.ss)-1)) + ylab("\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\n\nsamples")
  
```
  
## set heatmap size 
```{r heatmap_size, eval=TRUE,echo=F,fig.height=14,fig.width=20}
  heatmap.h<-5
  if(nrow(all.taxatabs.ss)>10) heatmap.h<-10
  if(nrow(all.taxatabs.ss)>100) heatmap.h<-50
  if(nrow(all.taxatabs.ss)>300) heatmap.h<-68
  if(nrow(all.taxatabs.ss)>400) heatmap.h<-88
  if(nrow(all.taxatabs.ss)>600) heatmap.h<-100
  if(nrow(all.taxatabs.ss)>700) heatmap.h<-120
  
  message("heatmap height=",heatmap.h," based on ",nrow(all.taxatabs.ss)," taxa")
  
  heatmap.w<-20
  #if(ncol(all.taxatabs.ss)>10) heatmap.w<-20
  #if(ncol(all.taxatabs.ss)>20) heatmap.w<-20
  if(ncol(all.taxatabs.ss)>70) heatmap.w<-30
  if(ncol(all.taxatabs.ss)>100) heatmap.w<-50
  
  message("heatmap width=",heatmap.w," based on ",ncol(all.taxatabs.ss)," samples")

```

## heatmaps 
```{r heatmaps, eval=TRUE,echo=F,fig.height=heatmap.h,fig.width=heatmap.w}
if(plot.heatmaps) if(!is.null(plotting.vars)){
  
  message("Making heatmaps")
  message("Note: Heatmaps will be larger in this html than if plotted alone")
  message("Secondary colour bars may not make sense if factors are overlapping, please check")
    
  if(!exists("values.to.plot")) values.to.plot<-c("nreads","ndxns","dxn")
  
  if(!exists("tidy.taxon.names")) tidy.taxon.names="order"
  
    for(i in 1:length(plotting.vars)){
          for(h in 1:length(values.to.plot)){
            
            plot.vals=T
            if(values.to.plot[h]=="dxn") plot.vals=F
            
            taxatab.heatmap(taxatab = all.taxatabs.ss,master_sheet = ms_ss,group.by=plotting.vars[i],
                values=values.to.plot[h],colour.bar=colour.bar,facetcol=facetcol,inc.values = plot.vals,
                current.grouping = sumrepsby,tidy.taxon.names = tidy.taxon.names)
        }
      }
}


```


## PCoA plots 
```{r pca plots, eval=TRUE,echo=F}
if(plot.pca) if(!is.null(plotting.vars)){
  message("Plotting PCoA plot with lines")
  
  if(!exists("plotting.vars.pca")) plotting.vars.pca<-plotting.vars
  
  for(j in 1:length(taxatabs.multilevels)){
    
    if(ncol(taxatabs.multilevels[[j]])>3) {
      
      message("The following plots are comparisons at ",names(taxatabs.multilevels)[j]," level")
      
    for(i in 1:length(plotting.vars.pca)){
      print(taxatab.pca.plot.col(taxatab = taxatabs.multilevels[[j]],ms_ss = ms_ss,grouping = sumrepsby,
                                 lines = show.pca.lines,longnames = F,
                                 shortnames = T,ellipse = T,factors = plotting.vars.pca[i],hidelegend = hidelegend.pca))
    }
    } else message("Only two samples at ", names(taxatabs.multilevels)[j]," level, skipping PCoA plot")
  }
  
 
} else message("plotting.vars=F...skipping pca plots")
```

## species accumulation curve 
```{r species accumulation curve, eval=TRUE,echo=F}
print(bas.plot.specaccum(taxatab = all.taxatabs.ss,xlabel = sumrepsby))
if(ncol(all.taxatabs.ss)>3) {
  print(bas.plot.specrich(taxatab = all.taxatabs.ss))
} else message("Only two samples in taxa table, skipping species richness plot")
```

## final table stats
```{r final table counts, eval=TRUE,echo=F}
#rank stats
print(stats.by.rank(taxatab = all.taxatabs.ss))

message("Reads total: ",(sum(colSums(all.taxatabs.ss[,-1,drop=F]))))

message("Reads per sample: 
        ")
print(summary(colSums(all.taxatabs.ss[,-1,drop=F])))

message("Mean reads per sample: 
        ")
print(mean(colSums(all.taxatabs.ss[,-1,drop=F])))

message("Standard deviation of mean reads per sample: 
        ")
print(sd(colSums(all.taxatabs.ss[,-1,drop=F])))

#summary
DT::datatable(summary.dxns.by.taxon(all.taxatabs.ss), rownames = FALSE, options = list(pageLength = 50, scrollX=T))

# final taxa list
#print(all.taxatabs.ss$taxon)
```

## qplot of % of detections that have x reads
```{r qplot final table, eval=TRUE,echo=F}
# qplots
plot1<-qplot.taxatab(all.taxatabs.ss,rm.nohits = T,rm.NA = T,specReadCount = filter_dxn)
print(plot1)
```

## combine counts and taxalists
```{r step counts, eval=TRUE,echo=F}
#counts
message("Note that detections/samples after sumreps will be fewer because reps are joined")

all.counts<-do.call(rbind,lapply(all.stats, `[[`, 1))
all.counts$step<-"not_assigned"
for(i in 1:length(all.stats)){
  all.counts$step[i]<-names(all.stats[[i]])[1]  
}
all.counts$step<-gsub("_counts","",all.counts$step)

print(all.counts)

#taxalists
message("The taxalist output needs more work, depends on collapsed taxa, need to split nto 2 tables, pre and post collapse.
        Running pre-aggregate_by_xLevel only. Some clever way to plot this?")

taxaSteps<-data.frame(start=all.stats[[1]][2][[1]])

#pre-collapse
namesSteps<-"names"
for(i in 1:length(all.stats)){
  namesSteps[i]<-names(all.stats[[i]])[2]  
}

taxaSteps2<-taxaSteps
for(i in 2:(grep("aggregate_by_xLevel_taxa",namesSteps)-1)){
  taxaSteps2<-merge(taxaSteps2, all.stats[[i]][2][[1]][,1:2],by.x = "start.taxon",by.y = "taxon",all.x = T)
  colnames(taxaSteps2)[length(colnames(taxaSteps2))]<-namesSteps[i]
}

DT::datatable(taxaSteps2, rownames = FALSE, options = list(pageLength = 10, scrollX=T,scrollY=T,fixedColumns = list(leftColumns = 1, rightColumns = 0)),extensions = c('FixedColumns'))

```

## plot counts
```{r plot counts, eval=TRUE,echo=F}
message("Note that detections/samples after sumreps will be fewer because reps are joined")

all.counts$step <- factor(all.counts$step,levels = all.counts$step)

print(all.counts)

stepPlot<-function(all.counts,type="reads"){
  ggplot(all.counts,aes(x=step,y=all.counts[,type],label = all.counts[,type]))+geom_bar(stat = "identity") +
   theme(axis.text.x = element_text(angle = 45,hjust=1))+
   theme(axis.line = element_line(colour = "black"))+
   theme(panel.background = element_blank())+
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   scale_y_continuous(labels = scales::comma)+
   ylab(type)
}

stepPlot(all.counts,type="reads")
stepPlot(all.counts,type="detections")
stepPlot(all.counts,type="taxa")
stepPlot(all.counts,type="samples")

#colour plot for reads, coloured by rank
for(i in 1:length(all.stats)){
  all.stats[[i]][[2]]$step<-as.character(all.counts$step[i])
}
all.stats.df<-do.call(rbind,lapply(all.stats,'[[',2))
all.stats.df$step <- factor(all.stats.df$step,levels = unique(all.stats.df$step))
print(ggplot(all.stats.df,aes(x=step,y=total.reads,fill=rank))+geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45,hjust=1))+
  theme(axis.line = element_line(colour = "black"))+
  theme(panel.background = element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_continuous(labels = scales::comma)+
  ylab("reads"))

```

## write final table
```{r write table, eval=TRUE,echo=F}
if(!is.null(taxatab.out)) write.table(all.taxatabs.ss,taxatab.out,quote = F, row.names = F,sep = "\t")
```

## make final krona plot
```{r write krona plot, eval=TRUE,echo=F}
if(!is.null(krona.out)) bas.krona.plot(taxatable = all.taxatabs.ss,out = krona.out)
```